```{r, include=FALSE}
source("custom_functions.R")
library(flextable)
library(officer)

```

---
title: "Exercise sheet 6: BLAST"
---

---------------------------------

# Exercise 1

You are given accession number \texttt{NM\_000667.3}. Use the [BLAST web server](http://blast.ncbi.nlm.nih.gov/Blast.cgi) to find
out about the gene that belongs to this acccession number (choose nucleotide blast, and the database RefSeq RNA).


#### {.tabset}

##### Question 1A

Which gene is it, and in which organism?

##### Solution

Gene: Alcohol Dehydrogenase 1A

Organism: *Homo sapiens* (human)

#### {-}

#### {.tabset}

##### Question 1B

Which other organisms does it seem to be highly conserved in?

##### Solution

- *Gorilla gorilla*: gorilla
- *Pan troglodytes*: common chimpanzee
- *Pan paniscus*: bonobo
- *Nomascus leucogenys*: northern white-cheeked gibbon
- *Cebus capucinus*: white-headed capuchin

Many more...

#### {-}

---------------------------------


# Exercise 2

You are given the sequences $A=\texttt{TAKTQHTSLPL}$ and $B=\texttt{TGTCTTCTGGGTTCAGCAAA}$ which stem from the same patient and are supposedly from the same gene. 

#### {.tabset}

##### Question 2A

What kind of sequences are these (DNA, RNA, Protein)? 

##### Solution 

$\texttt{A}$ is a protein, $\texttt{B}$ is a DNA sequence.

#### {-}

---------------------------------

#### {.tabset}

##### Question 2B

If you BLAST these sequences, for which sequence do you expect a lower *E-value*?

##### Solution

Even though sequence $\texttt{A}$ is shorter, as a sequence of amino acids it holds more information and should be more unique than sequence $\texttt{B}$. Hence the E-value for $\texttt{A}$ should be smaller.

#### {-}

---------------------------------

#### {.tabset}

##### Question 2C

BLAST both these sequences. Which gene do the sequences come from? Which BLAST result gave you a lower *E-value* and why? (use db: nucleotide collection and non-redundant protein sequences)

##### Solution

For sequence $\texttt{A}$ one needs to use blastp for sequence $\texttt{B}$ one needs to use blastn. Result is Forkhead box protein FBXW10. Sequence $\texttt{A}$ gives an E-value of $0.27$. $\texttt{B}$ gives $E=0.76$ (these numbers can change due to changes in the database, the important message is that $\texttt{A}$ < $\texttt{B}$ in this case). The reason for this is stated in the answer to the last question.


#### {-}

---------------------------------

# Exercise 3

You are given a nucleotide query sequence $q$ =  $\texttt{ATAC}$, and a nucleotide database sequence $s$ = $\texttt{ATAAAACGGGGGG}$. The word-size $k=2$. Use a simple scoring scheme that assigns a score of $2$ for a match and a score of $-1$ for a mismatch.

#### {.tabset}

##### Question 3A

Generate all $k$-length words of the query sequence.

##### Solution

- $w_1=AT$
- $w_2=TA$
- $w_3=AC$

#### {-}

---------------------------------

#### {.tabset}

##### Question 3B

List all possible words for the first $k$-length word (AT) that have a score of at least $T_1=1$.

##### Solution

- $s(AA)=1$
- $s(AC)=1$
- $s(AG)=1$
- $s(AT)=4$
- $s(CT)=1$
- $s(GT)=1$
- $s(TT)=1$

#### {-}

---------------------------------

#### {.tabset}

##### Question 3C

Scan the database for exact matches for the words from the question 3B.

##### Solution

*AA* at position 2,3,4. *AC* at position 5, *AT* at position 0.

#### {-}

---------------------------------

#### {.tabset}

##### Question 3D

Extend the exact matches that you found in the question 3C to the left/right and report all MSPs with a score greater than $4$.

##### Solution

AA:
```
Pos: 2              ATA
                    |||
                    AAA     with score 3
```
```
Pos: 3              ATAC
                    ||||
                    AAAC    with score 5
```
```
Pos: 4              AT
                    ||
                    AA    with score 1
```

AT:
```
Pos: 0              ATA
                    |||
                    AAA     with score 6
```

AC:
```
Pos: 5              A
                    |||
                    AAA     with score 6
```

MSPs start in the template at index 0 and 3.

---------------------------------

#### {.tabset}

##### Question 3E

What happens if we vary the parameters *k* and T_1$?

##### Solution

- Higher $T_1$, $k$: - faster (less seeds), - less sensitive (some hits will be missed)
- Lower $T_1$, $k$: - slower (more seeds), - more sensitive (less hits will be missed)

#### {-}

---------------------------------

```{r, include=knitr::is_html_output(), echo=F}
knitr::asis_output('Download: <a href="exercise-sheet-5.pdf" title="Download PDF Version">PDF Version</a> of this page.')
```

