```{r, include=FALSE}
source("custom_functions.R")
library(flextable)
library(officer)

```


---
title: "Exercise sheet 3: T-Coffee"
---

# Exercise 1

You are given the sequences $a$, $b$ and $c$

$$
a = CACCGG\\
b = ACCAAG\\
c = AACACC\\
$$

The pairwise optimal alignments $A(x, y)$ of the set of sequences $S$ were calculated as:

```
a: CACCG_G          a: __CACCGG     b: ACCAAG
    |||: |               ||||          |:||::
b: _ACCAAG          c: AACACC__     c: AACACC
```

#### {.tabset}

##### Question 1A

::: {.question data-latex=""}

Calculate the primary library ($L$)
:::


##### Formulae

::: {.answer data-latex=""}

init: $L^{x,y}_{i,j}=0$

$\forall$ alignments $A$ of sequences $x$ and $y$ of the set $S$. 

$weight(A) = \dfrac{number\ of\ matches}{min(len(x), len(y))} * 100$

$\forall$ aligned positions $i,j$ with $1 \leq i \leq len(x)$ and  $1 \leq j \leq len(y)$ 

$L^{x,y}_{i,j} = L^{x,y}_{i,j} + weight(A)$


:::


##### Solution

::: {.answer data-latex=""}

$L^{a,b}_{2,1}=L^{a,b}_{3,2}=L^{a,b}_{4,3}=L^{a,b}_{6,6}=100*\frac{4}{6}=67$ and all other $L^{a,b}_{i,j}=0$

$L^{a,c}_{1,3}=L^{a,c}_{2,4}=L^{a,c}_{3,5}=L^{a,c}_{4,6}=100*\frac{4}{6}=67$ and all other $L^{a,c}_{i,j}=0$

$L^{b,c}_{1,1}=L^{b,c}_{3,3}=L^{b,c}_{4,4}=100*\frac{3}{6}=50$ and all other $L^{b,c}_{i,j}=0$
:::
#### {-}

#### {.tabset}

##### Question 1B

::: {.question data-latex=""}

Calculate the extended library ($EL$)
:::

##### Formulae

::: {.answer data-latex=""}


$EL^{x,y}_{i,j}= L^{x,y}_{i,j}  + \sum_{z \in S \setminus \{\ x, y\}\ } \sum_{1 \leq k \leq len(z)}  min(L^{x,z}_{i,k}, L^{z,y}_{k,j})$

:::

##### Solution

::: {.answer data-latex=""}

The original Library doesn't change as there are no edges enforcing certain connections. Hence

$EL^{x,y}_{i,j}= L^{x,y}_{i,j} \qquad \forall L^{x,y}_{i,j}\neq 0$

and the following weights are added:

```
a: CACCG_G     
    |||: |    
b: _ACCAAG   
    |:||::
c:  AACACC
    * *
```

$EL^{a,b}_{1,3}=EL^{a,b}_{2,4}=50$

```
a: __CACCGG     
     ||||       
c: AACACC__     
   |:||::
b: ACCAAG
     **
```

$EL^{a,c}_{2,1}=EL^{a,c}_{4,3}=50$


```
b:    ACCAAG
      |||: |    
a:   CACCG_G     
     |||| 
c: AACACC
      ***
```

$EL^{b,c}_{1,4}=EL^{b,c}_{2,5}=EL^{b,c}_{3,6}=67$

:::
#### {-}

#### {.tabset}

##### Question 1C


Realign the sequences $b$ and $c$ using $EL$ for scoring and gap costs and mismatch costs of 0


##### Formulae

::: {.answer data-latex=""}

\begin{eqnarray*}
i &\in& [1,|x|] \\
j &\in& [1,|y|] \\
L(0,0) &=& 0 \\
L(i,0) &=& w(x_i,-)*i \quad \text{ or } = L(i-1,0)+w(x_i,-)\\
L(0,j) &=& w(-,y_j)*j \quad \text{ or } = L(0,j-1)+w(-,y_j)\\
L(i,j) &=& max \begin{cases}
	L(i-1,j)+w(x_i,-)  \\
	L(i,j-1)+w(-,y_j)  \\
	L(i-1,j-1)+w(x_i,y_j) 
	\end{cases}\\
w(x_i,y_j) &=& \begin{cases}
 EL^{x, y}_{i, j} & \text{match-score if } (x_i=y_j)\\
 0 & \text{insert/deletion-score if } (x_i = - \lor  y_j =-)\\
 0 & \text{mismatch-score else } (x_i\neq y_j) \\
	\end{cases} \\
\end{eqnarray*}

:::


##### Solution

```{r, include=FALSE}


onec <- data.frame(a = c("-", "A", "A", "C", "A", "C", "C"),
                   b = c(0, 0, 0, 0, 0, 0, 0),
                   c = c(0, 50, 50, 50, 67, 67, 67),
                   d = c(0, 50, 50, 50, 67, 133, 133),
                   e = c(0, 50, 50, 100, 100, 133, 200),
                   f = c(0, 50, 50, 100, 150, 150, 200),
                   g = c(0, 50, 50, 100, 150, 150, 200),
                   h = c(0, 50, 50, 100, 150, 150, 0))
names(onec) <- c("-1",  "-", "A1",   "C2", "C3",  "A4",   "A5",   "G6")

```
```{r, results="asis", include=knitr::is_html_output(), echo=FALSE}
onec_ft <- flextable(onec)
onec_ft <- custom_theme(onec_ft)
index_replace(onec_ft)

```
```{r, results="asis", include=knitr::is_latex_output(), echo=FALSE}
knitr::kable(onec)
```

#### {-}


#### {.tabset}

##### Question 1D

Do the other alignments $a$-$b$ and $a$-$c$ change? Provide arguments, without calculating new alignments.

##### Solution

No. The newly added alignment scores in $EL$ represent edges that are incompatible with the
current best alignments and can not score higher.

#### {-}

#### {.tabset}


##### Question 1E

Sketch a Guide Tree (either Sketch or Newick format)

##### Solution

Newick: $((a,c),b)$  or $((a,b),c)$

#### {-}

#### {.tabset}

##### Question 1F

Perform a progressive alignment by aligning sequence $b$ to the already existing alignment $A(a, c)$.
To score a match between $b$ and $A(a, c)$ use the sum $EL^{a,b} + EL^{b,c}$ with the correct indices.
Show the resulting multiple sequence alignment.

##### Solution

| -<br />- | -<br />- | -<br />A  | -<br />A  | C<br />C   | A<br />A   | C<br />C   | C<br />C   | G<br />-   | G<br />-   |
|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|
| __-__ | 0 | 0  | 0  | 0   | 0   | 0   | 0   | 0   | 0   |
| __A__ | 0 | 50 | 50 | 50  | 133 | 133 | 133 | 133 | 133 |
| __C__ | 0 | 50 | 50 | 50  | 133 | 267 | 267 | 267 | 267 |
| __C__ | 0 | 50 | 50 | 150 | 150 | 267 | 400 | 400 | 400 |
| __A__ | 0 | 50 | 50 | 150 | 250 | 267 | 400 | 400 | 400 |
| __A__ | 0 | 50 | 50 | 150 | 250 | 267 | 400 | 400 | 400 |
| __G__ | 0 | 50 | 50 | 150 | 250 | 267 | 400 | 400 | 467 |


##### Solution

```{r, include=FALSE}


onec <- data.frame(a = c("-", "A", "A", "C", "A", "C", "C"),
                   b = c(0, 0, 0, 0, 0, 0, 0),
                   c = c(0, 50, 50, 50, 67, 67, 67),
                   d = c(0, 50, 50, 50, 67, 133, 133),
                   e = c(0, 50, 50, 100, 100, 133, 200),
                   f = c(0, 50, 50, 100, 150, 150, 200),
                   g = c(0, 50, 50, 100, 150, 150, 200),
                   h = c(0, 50, 50, 100, 150, 150, 0))
names(onec) <- c("-1",  "-", "A1",   "C2", "C3",  "A4",   "A5",   "G6")

```
```{r, results="asis", include=knitr::is_html_output(), echo=FALSE}
onec_ft <- flextable(onec)
header = c("-", "-", "-", "-", "-", "-", "-", "-")
onec_ft <- custom_theme2headers(onec_ft, header)
index_replace(onec_ft)

```
```{r, results="asis", include=knitr::is_latex_output(), echo=FALSE}
knitr::kable(onec)
```

#### {-}

#### {-}

-------------------------------------------

```{r, include=knitr::is_html_output(), echo=F}
knitr::asis_output('Download: <a href="exercise-sheet-3.pdf" title="Download PDF Version">PDF Version</a> of this page.')
```
